\papertitle

### 2012-10-13 manuscript.tex
Edit after merge

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABSTRACT
\begin{abstract}
The power, size, and weight constraints of micro air vehicles (MAVs) limit 
their on-board sensing and computational resources. Ground vehicles have 
less mobility than MAVs, but relaxed size constraints and typically more 
computing power. These specializations present many opportunities for cooperation. In this work, we 
demonstrate cooperative target-seeking between a 13 gram ornithopter MAV 
and a lightweight ground station using computer vision. We develop models for the ornithopter, ground station, and cooperative system dynamics. We determine model parameters of the real systems through experimental system identification. Finally, we verify those models using experiments on narrow passage traversal, and arrive at a cooperative system model which accurately predicts the backwards-reachable region for successfully negotiating ornithopther flight through narrow openings.

We also introduce a new ornithopter MAV, the 13 gram H$^2$Bird. It 
features clap and fling wings, improves upon a previous design by 
utilizing a carbon fiber airframe, tail rotor, and elevator, and carries a 
payload of 2.8 grams. We augment the ornithopter's built-in 
gyroscope-based control with a lightweight ground station design, which has power and weight 
requirements appropriate for deployment on a millirobotic ground vehicle. The ground 
station provides heading estimates to the ornithoper by running a 
real-time motion tracking algorithm on a live video stream. This allows 
the ornithopter to reliably negotiate flight through narrow openings. We demonstrate the performance of the location estimation and navigation  algorithms.
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ACM METADATA

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\begin{figure}[!tb]
\centering
\includegraphics[width=\linewidth]{figures/h2bird.png}
\caption{H$^2$Bird ornithopter MAV}
\label{fig:h2bird}
\end{figure}
Power, size, and weight constraints on micro air vehicles (MAV) 
significantly limit sensing their and computational capabilities. They can execute only the most rudimentary
control and state estimation algorithms in real-time. 
The constraints on flapping-wing MAVs are even more restrictive. 
Though they exhibit demonstrate safety and noise profiles superior to those
of rotor craft MAVs, this comes at the expense of reduced payload weight.

The non-linear
flight dynamics exhibited of flapping wing flight compounding constraints~\cite{humbert:dipteran2}\cite{cheng:tandrdamping}. 
Much past research on flapping wing MAVs focused on overcoming the 
difficulties associated with controlling and modeling flapping-wing vehicles in the face of this
non-linearity\cite{floreano:flying}. Common approaches to simplify these 
dynamic models include averaging over the wingbeat period~\cite{Schenato:flightcontrol}
\cite{cheng:tandrdamping}, and using linear, low-dimensional models to predict flight behavior~\cite{Shigeoka:ornithopter}. These advances lead to the
development of some high-performance platforms, such as the sub-15 gram DelFly 
described by deCroon et al.~\cite{delfly:design}.

Many control laws that previous researchers developed from
dynamic models of flapping-winged flight, such as periodic forcing and other
control methods described %TODO you need to use real names, it's not kosher to refer to citation numbers directly
by~\cite{doman:dynamics}\cite{khan:longitudinal_control}\cite{leonard:averaging}, 
require powerful on-board computing. In this work, we develop a simpler, less
computationally intensive method of control for negotiating narrow passages, by exploiting cooperation between specialized robotic agents.

In contrast to flight control, there is considerably less previous work on guidance and of flapping-wing MAVs. 
Baek demonstrated altitude control with an external camera~\cite{baek:altitude}. 
de Croon et al. demonstrated obstacle avoidance with an onboard camera and offboard 
processing~\cite{delfly:avoid}. For sub-20 gram ornithopters, Baek et al. 
\cite{baek:tracking} provides one of the few examples of completely autonomous target seeking for vehicles at this scale.

There also exists a body of work on the guidance and control of fixed- and rotary-wing 
MAV platforms, but many of these methods make extensive use of GPS or motion capture systems
~\cite{kingston:timeattitude}\cite{kanade:3dvision}. Indoors, using
GPS to estimate the pose of an MAV is impossible, and motion capture
requires extensive and stationary modification of the environment. The power
and payload limits on flapping wing MAVs make complex on-board sensing difficult, so multi-agent
cooperative control schemes are apt solutions.

\cite{Jung1998Range}\cite{Hyams1999Cooperative}\cite{Mehta2006Adaptive}
\cite{Rudol2008Micro}\cite{Luo2011Air}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Robotic and Vision Platforms}
%----------------------------------------------------------------------------%
\subsection{H$^2$Bird Ornithoper}
% humphrey
To explore these cooperative control concepts, we designed a new flapping-wing MAV, known as the H$^2$Bird (Fig.~\ref{h2bird}). Built around the Silverlit i-Bird 
RC flyer power train and clap-fling 
wings\footnote{\raggedright Silverlit Toys Manufactory Ltd.: i-Bird RC Flyer
\href{http://www.silverlit-flyingclub.com/wingsmaster/}
     {http://www.silverlit-flyingclub.com/wingsmaster/}}, 
the H$^2$Bird has a wingspan of 26.5 cm and a flight weight of 13 grams. A 
tail propeller and servo-controlled tail elevator provide control in the yaw 
and pitch axes (Fig.~\ref{h2Bird_axes}. The on-board ImageProc 2.4\footnote{ImageProc 2.4: \\
\href{https://github.com/biomimetics/imageproc\_pcb}
     {https://github.com/biomimetics/imageproc\_pcb}} 
controller consists of a 40 MIPS microprocessor, 6 DOF IMU, 
IEEE 802.15.4 radio, VGA camera, and motor drivers, all powered by a 90 mAh 
lithium polymer battery. In routine flight, the H$^2$Bird averages 1.2 m/s 
ground speed, and operates for approximately 10 minutes.

%----------------------------------------------------------------------------%
\subsection{Ground Station Computer Vision Platform}
% ryan
Cooperative behavior is a promising strategy for autonomous navigation of 
unexplored and unstructured environments. However, to operate in these environments,
a cooperative system must be fully mobile, not tied to a static ground station or 
motion capture laboratory. To demonstrate the feasibility of this approach, 
we designed our ground station to conform to the power and weight 
constraints of highly mobile millirobotic ground vehicles, such as 
OctoRoACH, which can traverse rough terrains with speed and agility~\cite{Pullin2012Dynamic}.

Rather than power-intensive PC processors, we used the ARM-based 
BeagleBoard single-board computer as reference computation platform. The
BeagleBoard consumes approximately one watt while running our computer vision 
and control algorithms. It uses the same processor as popular sub-10 gram 
processor modules---such as the Gumstix Overo and LogicPD Torpedo---which are 
light enough for deployment on millirobotic ground vehicles. We pair the 
BeagleBoard with an off-the-shelf consumer USB web camera, for similar image 
quality and resolution to the miniature camera modules available at 
millirobot scale. The BeagleBoard communicates with the H$^2$Bird with a
USB radio module which implements the IEEE 802.15.4 wireless standard.
%TODO footnotes: Gumstix, LogidPD

@@

We implement attitude estimation and control of the H$^2$Bird on-board to achieve 
high sample-rate attitude control. We estimate the vehicle pose by
integrating the IMU, and use a proportional-integral-derivative (PID) 
controller to feed these estimates back to the attitude control flight surfaces (Figs.\ref{process_flow} and~\ref{block_diagram}).

The high pitch angles flapping-wing MAVs regularly operate at distinguish them 
from other MAVs. Baek demonstrated control on a similar vehicle using PID and an Euler angle
parameterization of orientation~\cite{baek:tracking}. However, Euler angle parameterizations of vehicle orientation 
suffer from singularities around high pitch angles. So we instead 
represent the vehicle orientation with quaternions. An additional benefit of 
the quaternion representation is its ability to represent relative body and 
world coordinate angular displacements compactly with quaternion multiplication.

@@

%----------------------------------------------------------------------------%
\subsection{Ground Station Pose Estimation}
The ground station performs simple two-dimensional pose estimation on 
H$^2$Bird using a particle filter-based motion tracking algorithm.
The produced pose estimate is the feedback input to the cooperative visual servoing feedback 
loop (Sec. \ref{sec:visual_servoing_concept}). We limit our pose estimation 
approach to monocular two-dimensional object tracking in pixel space. This 
allows us to perform real-time video tracking using the modest computational 
resources available on the ground station, and similarly, millirobotic ground 
vehicles.

Our pose estimation algorithm tracks H$^2$Bird's motion in the frame using a particle filter ~\cite{thrun2005probabilistic}. 
We initialize the particles uniformly across the frame before tracking begins. 
To improve numerical stability, we normalize the weights of all particles on 
each iteration, and only resample the particle population when the mean particle 
weight falls below a pre-determined threshold.

%............................................................................%
\subsubsection{Motion Model}
Our motion model (Eq.~\ref{eq:motion_model}) is a simple Gaussian transition 
centered around each particle's current position, augmented with an 
$\epsilon$-random uniform sampling strategy. In Eq.~\ref{eq:motion_model}, 
$x^{[i]}_t$ is the position of the $i^\text{th}$ particle in pixel space at time $t$, $\Sigma$ 
is a chosen covariance matrix, and $b$ is vector of the bounds of the pixel space.
\begin{equation}
\label{eq:motion_model}
x^{[i]}_t \sim \begin{cases}
\mathcal{N}(x^{[i]}_{t-1}, \Sigma), &\text{with probability } 1-\epsilon \\
\mathcal{U}(0,b),& \text{with probability } \epsilon \\
\end{cases}\\
\end{equation}
In the experiments we present in this work, we used a diagonal covariance matrix 
with a uniform variance of 256, and an $\epsilon$ of 0.3. We find the Gaussian 
transition model is sufficient for tracking motion in which frame-to-frame 
position changes are relatively small. The $\epsilon$-random uniform 
transitions increase tracking robustness, by forcing the filter to always 
sample broadly across of the pixel space. This helps prevent overconfidence and collapse, 
and decreases reacquisition delay.

%............................................................................%
\subsubsection{Emission Model}
Our emission model (Eq.~\ref{eq:emission_model}) implements a sampling
approach to motion tracking via na\"{i}ve background subtraction, e.g. frame differencing[CITATION].
On each iteration, the system provides the emission model with $f_0$, the frame representing 
the background and $f_t$, the most recent frame captured. The model assigns 
a weight $w^{[i]}_t$ to each particle $x^{[i]}_t$, proportional to the color 
distance in RGB space between the pixels of $f_0$ and $f_t$ at the location 
$x^{[i]}_t$.
\begin{equation}
\label{eq:emission_model}
w^{[i]}_t = 1 + \norm{f_0(x^{[i]}_t)-f_t(x^{[i]}_t)}^2
\end{equation} 
Hence, particles are more likely to be resampled in regions of 
the image which are dissimilar in color from the background.

During experiments presented in this work, we assumed that neither the bird, 
nor any other moving object, were visible in the frame on system 
start. We saved the first frame captured by the system, and provided it as 
the background frame $f_0$ for every iteration of the particle filter. This choice of $f_0$ 
resulted in a sampling approach to na\"{i}ve background subtraction.

Na\"{i}ve background subtraction is known to be brittle to changes in the scene over 
time[CITATION]. However, we find it very efficient and sufficiently reliable for the 
short time spans necessary for H$^2$Bird to successfully negotiate narrow 
openings, which are typically less than five seconds. Substituting $f_0$ 
with a more sophisticated background extraction would retrain the underlying motion tracking features of the filter, while providing it with a more robust way of differentiating subject motion from the background.

%----------------------------------------------------------------------------%
\subsection{Cooperative Visual Servoing}
\label{sec:visual_servoing_concept}
The ground station guides the H$^2$Bird by commanding changes to its
reference attitude, as governed by its on-board attitude controllers (Figs.\ref{process_flow} and~\ref{block_diagram}). 
With the estimate of 2D vehicle pose from computer vision, we calculate an altitude
error and horizontal error. The ground station then commands a reference attitude to 
the H$^2$Bird pitch and yaw. according to a PID control law. We apply a fixed-lag window the PID controller integrator term 
to prevent wind-up while we launch the ornithopter. We tuned the proportional and derivative terms of this outer controller 
to dampen the overshoot and drift cause by the cooperative system latency.

\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{figures/block_diagrams.pdf}
\caption{Overall block diagram of the cooperative control system.}
\label{fig:block_diagram}
\end{figure}

@@

We approximate the yaw dynamics of the ornithopter with a one-dimensional model to limit complexity and because it matches our physical implementation. Likewise, we choose the reference heading angle as input to the model and model the dynamics of vehicle and on-board PID controller as one system, which outputs the physical heading angle. We show the resulting model form in Equation~\ref{eq:plain_model}.

\begin{equation}
\label{eq:plain_model}
H(s) = \frac{K}{1 + Rs}
\end{equation}

\begin{equation}
\label{eq:dubin_car}
\begin{aligned}
& \dot{x} = u_{s}sin(\theta)\\
& \dot{y} = u_{s}cos(\theta)\\
& \dot{\theta} = u\\
& u \in U=[-\tan{\phi_{max}},\tan{\phi_{max}}]
\end{aligned}
\end{equation}

To model the translational position of the ornithopter, we use a form of Dubin's car model (Eq. ~\ref{eq:dubin_car})~\cite{lavalle:planning}, where $\phi_{max}$ is the maximum steering angle and $u_{s}$ is the forward speed. We modify this model to treat the angular position $\theta$ as an input rather than a state. The overall motion of the system can be modeled as shown in Equation~\ref{eq:total_model_plain}, where $\theta$ is the time domain output of (\ref{eq:plain_model}) with the reference heading angle as the input.

Our form uses the angular position, $\theta$ as an input rather than a state. We model the overall motion of the system as shown in Equation~\ref{eq:total_model_plain}, where $\theta$ is the time domain output of (Eq. \ref{eq:plain_model}) with the reference heading angle as the input.

@@

Our model of the ground station comprises models of both the camera and the
pose estimation computation. The camera model contains both a
latency, which is in series with the path between the ``camera'' and ``ornithopter dynamics''
 blocks in Fig.~\ref{fig:block_diagram}, and a model of the camera's viewing angle constraints.

 We model the 
viewing angle of the camera as a pixel map from one edge of a
63$^{\circ}$ viewing triangle to the other. The camera has an image width of
640 pixels. We calculate the desired heading angle in degrees for a given distance from the
camera according to Equation~\ref{eq:desired_heading}, where $d$
is the distance between the ornithopter and the camera, and $x$ is the horizontal
position of the robot relative to the center of the window in meters. This
relation reflects the fact that the camera cannot determine the depth of the
robot. One example of why this is a problem is that if the robot stays at a
constant horizontal distance from the window, but moves closer and closer to
the camera, the directed heading steadily increases even though the robot has
not actually moved farther from the window. This creates tracking problems in
areas close to the camera.

\begin{equation} \label{eq:desired_heading} \text{Desired heading} =
\frac{640\text{px}}{2*\text{d}*\tan{\frac{63^{\circ}}{2}}}*\frac{63^{\circ}}{640\text{px}}*\text{x}
\end{equation}

We model the pose estimation computation by the by the latency of running 
the particle filter, calculating a new reference attitude, and sending to the ornithopter via radio.

@@

%............................................................................%
\subsubsection{Ornithoper Attitude Control}

We determined the turning speed and turning radius of the H$^2$Bird
experimentally, by measuring the response of the system to a step input of a
clockwise 90$^{\circ}$ turn. We measured the response of the system by recording
the attitude estimates calculated by the ornithopter using its on-board
gyroscope. We then used
MATLAB\footnote{The MathWorks, Inc. Matlab:
\href{http://www.mathworks.com/products/matlab/}
     {http://www.mathworks.com/products/matlab/}} to fit a
simple low-order model to the step response, as discussed in Sec.~\ref{sec:model_attitude}.

To measure the capture latency of the camera, we wrote a simple test 
application which changes a large colored rectangle on the ground station's screen, then uses the camera to detect that change. By noting the value of the system timer just before we change the screen, and just after the camera detects that change, the application arrives at an estimate of the video capture pipeline latency. We averaged these result over several hundred rapidly-executed trials to determine the capture latency we used to model the system..

@@

We tested our cooperative visual servoing system with a simple target seeking
experiment: the H$^2$Bird must fly through a specified obstacle, in this case,
a wooden frame we used to simulate a window. As the H$^2$Bird cannot detect the
target on its own, we used the cooperative control with the ground station, which to provided remote guidance to H$^2$Bird to allow it to navigate to the window.
We used a motion capture system to record 
ground truth data. We present a sketch of this experimental setup in
Fig.~\ref{fig:experiment_cartoon}.

% TODO clarify or roll into paragraph
Each trial included the following steps:
\begin{enumerate}
\item The window is identified by the user. 
\item We release H$^2$Bird by hand, in view of the camera at the desired starting grid point.
\item The H$^2$Bird attempts to fly through the window with, with remote guidance from
the ground station.
\end{enumerate}

In total, we ran 80 trials over a variety of starting positions. Our goal for the experiments was to verify the model described in
Secttion~\ref{sec:system_model}, so we conducted 60 of the 80 trials in a grid
along the edges of the camera view plane, and in 0.05 meter increments from the
window, in an effort to determine the success rate on the edges of the testing space. We
conducted 20 additional trials directly in front of the camera to determine
the success rate near the center of viewing plane. After every 5 trials,
we replaced the ornithopter's battery with a fully charged one.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABSTRACT
\begin{abstract}
The power, size, and weight constraints of micro air vehicles (MAVs) limit 
their on-board sensing and computational resources. Ground vehicles have 
less mobility than MAVs, but relaxed size constraints, and typically more 
computing power. These specializations present many opportunities for cooperation. In this work, we 
demonstrate cooperative target-seeking between a 13 gram ornithopter MAV 
and a lightweight ground station using computer vision. We demonstrate the performance of the location estimation and navigation algorithms. We develop models for the ornithopter, ground station, and cooperative system dynamics. We determine model parameters of the real systems through experimental system identification. Finally, we verify those models using experiments on narrow passage traversal, and arrive at a cooperative system model which accurately predicts the backwards-reachable region for successfully negotiating ornithopther flight through narrow openings.

We also introduce a new ornithopter MAV, the 13 gram H$^2$Bird. It 
features clap and fling wings, improves upon a previous design by 
utilizing a carbon fiber airframe, tail rotor, and elevator, and carries a 
payload of 2.8 grams. We augment the ornithopter's built-in 
gyroscope-based control with a lightweight ground station design, which has power and weight 
requirements appropriate for deployment on a millirobotic ground vehicle. The ground 
station provides heading estimates to the ornithoper by running a 
real-time motion tracking algorithm on a live video stream. This allows 
the ornithopter to reliably negotiate flight through narrow openings.
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ACM METADATA

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\begin{figure}[!tb]
\centering
\includegraphics[width=\linewidth]{figures/h2bird.png}
\caption{H$^2$Bird ornithopter MAV}
\label{fig:h2bird}
\end{figure}
Power, size, and weight constraints on micro air vehicles (MAV) 
significantly limit their sensing and computational capabilities. They can 
execute only the most rudimentary control and state estimation algorithms in 
real-time.  The constraints on flapping-wing MAVs are even more restrictive. 
Though they demonstrate safety and noise profiles superior to those of 
rotorcraft MAVs, this comes at the expense of reduced payload.

The non-linear dynamics flapping-wing flight compound 
these constraints~\cite{humbert:dipteran2}\cite{cheng:tandrdamping}. Much 
past research on flapping-wing MAVs focuses on overcoming the difficulties 
of controlling and modeling flapping-wing vehicles in the face of this
non-linearity\cite{floreano:flying}. Common approaches to simplify these 
dynamic models include averaging over the wing-beat 
period~\cite{Schenato:flightcontrol}\cite{cheng:tandrdamping}, and using 
linear, low-dimensional models to predict flight 
behavior~\cite{Shigeoka:ornithopter}. These advances lead to the development 
of high-performance platforms such as the sub-15 gram DelFly, described by 
deCroon et al.~\cite{delfly:design}.

Many control laws previous researchers developed from dynamic models of 
flapping-wing flight, such as periodic forcing and other control methods 
described by~\cite{doman:dynamics}\cite{khan:longitudinal_control}\cite{leonard:averaging}, 
require powerful on-board computing. In this work, we develop a simpler, 
less computationally-intensive method of control for negotiating narrow 
passages, by exploiting cooperation between specialized robotic agents.

In contrast to flight control, there is considerably less previous work on 
guidance and navigation of flapping-wing MAVs. Baek demonstrated altitude 
control with an external camera~\cite{baek:altitude}. de Croon et al. 
demonstrated obstacle avoidance with an onboard camera and offboard 
processing~\cite{delfly:avoid}. Baek et al. provide one of the few examples 
of completely autonomous target seeking for vehicles at the sub-20 gram 
scale~\cite{baek:tracking}.

There also exists a body of work on the guidance and control of fixed- and 
rotary-wing MAV platforms, but many of these methods make extensive use of 
GPS or motion capture systems~\cite{kingston:timeattitude}\cite{kanade:3dvision}. 
Indoors, using GPS to estimate the pose of an MAV is impossible, and motion 
capture requires extensive and stationary modification of the environment. 
The power and payload limits on flapping-wing MAVs make complex on-board 
sensing difficult, so multiagent cooperative control schemes are apt 
solutions.

%TODO Ryan's paragraph
\cite{Jung1998Range}\cite{Hyams1999Cooperative}\cite{Mehta2006Adaptive}
\cite{Rudol2008Micro}\cite{Luo2011Air}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Robotic and Vision Platforms}
%----------------------------------------------------------------------------%
\subsection{H$^2$Bird Ornithoper}
% humphrey
To explore these cooperative control concepts, we designed a new 
flapping-wing MAV, known as the H$^2$Bird (Fig.~\ref{fig:h2bird}). Built 
around the Silverlit i-Bird RC flyer power train and clap-fling wings\footnote{\raggedright Silverlit Toys Manufactory Ltd.: i-Bird RC Flyer
\href{http://www.silverlit-flyingclub.com/wingsmaster/}
     {http://www.silverlit-flyingclub.com/wingsmaster/}}, 
the H$^2$Bird has a wingspan of 26.5 cm and a flight weight of 13 grams. A 
tail propeller and servo-controlled tail elevator provide control in the yaw 
and pitch axes (Fig.~\ref{fig:h2Bird_axes}). The on-board ImageProc 2.4\footnote{ImageProc 2.4: \\
\href{https://github.com/biomimetics/imageproc\_pcb}
     {https://github.com/biomimetics/imageproc\_pcb}} 
controller includes of a 40 MIPS microprocessor, 6 DOF IMU, 
IEEE 802.15.4 radio, VGA camera, and motor drivers, all powered by a 90 mAh 
lithium polymer battery. In routine flight, the H$^2$Bird averages 1.2 m/s 
ground speed, and operates for approximately 10 minutes.

%----------------------------------------------------------------------------%
\subsection{Ground Station Computer Vision Platform}
% ryan
Cooperative behavior is a promising strategy for autonomous navigation in 
unexplored and unstructured environments. However, to operate in these 
environments, a cooperative system must be fully mobile, not tied to a 
static ground station or motion capture laboratory. To demonstrate the 
feasibility of this approach, we designed our ground station to conform to 
the power and weight constraints of highly mobile millirobotic ground 
vehicles, such as OctoRoACH, which can traverse rough terrains with speed and 
agility~\cite{Pullin2012Dynamic}.

Rather than power-intensive PC processors, we used the ARM-based BeagleBoard 
single-board computer as reference computation platform. The BeagleBoard 
consumes approximately one watt while running our computer vision and 
control algorithms. It uses the same processor as popular sub-10 gram 
processor modules--such as the Gumstix Overo and LogicPD Torpedo--which are 
light enough for deployment on millirobotic ground vehicles. We pair the 
BeagleBoard with an off-the-shelf consumer USB web camera, for similar image 
quality and resolution to the miniature camera modules available at 
millirobot scale. The BeagleBoard communicates with the H$^2$Bird with a 
USB radio module which implements the IEEE 802.15.4 wireless standard.
%TODO footnotes: Gumstix, LogidPD

@@

We implement attitude estimation and control of the H$^2$Bird on-board to 
achieve high sample-rate attitude control. We estimate the vehicle pose by
integrating the IMU, and use a proportional-integral-derivative (PID) 
controller to feed these estimates back to the attitude control flight 
surfaces (Figs.~\ref{fig:process_flow} and~\ref{fig:block_diagram}).

The high pitch angles at which flapping-wing MAVs regularly operate 
distinguish them from other MAVs. Baek demonstrated control on a similar 
vehicle using PID and an Euler angle parameterization of orientation~\cite{baek:tracking}. 
However, Euler angle parameterizations of vehicle orientation suffer from 
singularities around high pitch angles. So we instead represent the vehicle 
orientation with quaternions. The quaternion representation allows us to 
represent relative body and world coordinate angular displacements compactly 
with quaternion multiplication.

@@

%----------------------------------------------------------------------------%
\subsection{Ground Station Pose Estimation}
The ground station performs simple two-dimensional pose estimation on 
H$^2$Bird using a particle filter-based motion tracking algorithm. This pose 
estimate is the feedback input to the cooperative visual servoing feedback 
loop, which is represented as the outermost feedback paths on the block 
diagrams in Figs.~\ref{fig:process_flow} and~\ref{fig:block_diagram}. We 
limit our pose estimation approach to monocular two-dimensional object 
tracking in pixel space. This allows us to perform real-time video tracking 
using the modest computational resources available on the ground station, 
and similarly, millirobotic ground vehicles.

Our pose estimation algorithm tracks H$^2$Bird's motion in the frame using a 
particle filter~\cite{thrun2005probabilistic}. We initialize the particles 
uniformly across the frame before tracking begins. To improve numerical 
stability, we normalize the weights of all particles on each iteration, and 
only resample the particle population when the mean particle weight falls 
below a pre-determined threshold. 
%............................................................................%
\subsubsection{Motion Model}
Our motion model (Eq.~\ref{eq:motion_model}) is a simple Gaussian transition 
centered around each particle's current position, augmented with an 
$\epsilon$-random uniform sampling strategy. In Eq.~\ref{eq:motion_model}, 
$x^{[i]}_t$ is the position of the $i^\text{th}$ particle in pixel space at 
time $t$, $\Sigma$ is a chosen covariance matrix, and $b$ is a vector of the 
bounds of the pixel space.
\begin{equation}
\label{eq:motion_model}
x^{[i]}_t \sim \begin{cases}
\mathcal{N}(x^{[i]}_{t-1}, \Sigma), &\text{with probability } 1-\epsilon \\
\mathcal{U}(0,b),& \text{with probability } \epsilon \\
\end{cases}\\
\end{equation}
In the experiments we present in this work, we used a diagonal covariance matrix 
with a variance of 256, and an $\epsilon$ of 0.3. We find the Gaussian 
transition model is sufficient for tracking motion in which frame-to-frame 
position changes are relatively small. The $\epsilon$-random uniform 
transitions increase tracking robustness, by forcing the filter to always 
sample broadly across of the pixel space. This helps prevent overconfidence 
and collapse, and decreases reacquisition delay.

%............................................................................%
\subsubsection{Emission Model}
Our emission model (Eq.~\ref{eq:emission_model}) implements a sampling
approach to motion tracking via na\"{i}ve background subtraction, e.g. frame 
differencing[CITATION]. On each iteration, the system provides the emission 
model with $f_0$, the frame representing the background, and $f_t$, the most 
recent frame captured. The model assigns a weight $w^{[i]}_t$ to each 
particle $x^{[i]}_t$, proportional to the color distance in RGB space 
between the pixels of $f_0$ and $f_t$ at the location $x^{[i]}_t$.
\begin{equation}
\label{eq:emission_model}
w^{[i]}_t = 1 + \norm{f_0(x^{[i]}_t)-f_t(x^{[i]}_t)}^2
\end{equation} 
Hence, particles are more likely to be resampled in regions of 
the image which are dissimilar in color from the background.

During the experiments presented in this work, we assumed that neither the 
bird, nor any other moving object, were visible in the frame on system 
start. We saved the first frame captured by the system, and provided it as 
the background frame $f_0$ for every iteration of the particle filter. This 
choice of $f_0$ resulted in a sampling approach to na\"{i}ve background 
subtraction.

Na\"{i}ve background subtraction is brittle to changes in a scene over 
time [CITATION]. However, we find it very efficient, and sufficiently 
reliable for the short time spans--less than five seconds--necessary for 
H$^2$Bird to successfully negotiate narrow openings. Calculating $f_0$ 
using a more sophisticated background extraction would retrain the 
underlying motion tracking features of the filter, while providing the 
algorithm with a more robust way of differentiating subject motion from the 
background.

%----------------------------------------------------------------------------%
\subsection{Cooperative Visual Servoing}
\label{sec:visual_servoing_concept}
The ground station guides the H$^2$Bird by commanding changes to its 
reference attitude, which is in turn governed by its on-board attitude 
controllers (Figs.\ref{fig:process_flow} and~\ref{fig:block_diagram}). Using 
the estimate of 2D vehicle pose from the pose estimation algorithm, the 
ground station calculates an altitude error and horizontal error. The ground 
station then commands a reference attitude to the H$^2$Bird over radio, 
according to a PID control law. We apply a fixed-lag window the PID 
controller integral term to prevent wind-up while we launch the ornithopter. 
We tuned the proportional and derivative terms of this outer controller to 
dampen the overshoot and drift caused by the cooperative system latency.
\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{figures/block_diagrams.pdf}
\caption{Overall block diagram of the cooperative control system.}
\label{fig:block_diagram}
\end{figure}

@@

We approximate the yaw dynamics of the ornithopter with a one-dimensional 
model to limit complexity, and because it matches our physical 
implementation. Likewise, we choose the reference heading angle as input to 
the model, and model the dynamics of vehicle and on-board PID controller as 
one system, whose output is the physical heading angle. We show the 
resulting model form in Equation~\ref{eq:plain_model}.
\begin{equation}
\label{eq:plain_model}
H(s) = \frac{K}{1 + Rs}
\end{equation}
To model the translational position of the ornithopter, we use a form of 
Dubin's car model (Eq.~\ref{eq:dubin_car})~\cite{lavalle:planning}, where 
$\phi_{max}$ is the maximum steering angle and $u_{s}$ is the forward speed. 
\begin{equation}
\label{eq:dubin_car}
\begin{aligned}
& \dot{x} = u_{s}sin(\theta)\\
& \dot{y} = u_{s}cos(\theta)\\
& \dot{\theta} = u\\
& u \in U=[-\tan{\phi_{max}},\tan{\phi_{max}}]
\end{aligned}
\end{equation}
We modify this model to treat the angular position $\theta$ as an input 
rather than a state. The overall motion of the system can be modeled as shown 
in Equation~\ref{eq:dubin_car}, where $\theta$ is the time-domain output of 
Equation~\ref{eq:plain_model} with the reference heading angle as input.

@@

Our model of the ground station comprises models of both the camera and the
pose estimation computation. The camera model contains both a latency, which 
is in series with the path between the ``Camera'' and 
``Ornithopter Dynamics'' blocks in Figure~\ref{fig:block_diagram}, and a 
model of the camera's viewing angle constraints.

We model the viewing angle of the camera as a pixel map from one edge of a 
63$^{\circ}$ viewing triangle to the other. The camera has an image width of 
640 pixels. We calculate the desired heading angle in degrees, for a given 
distance from the camera, according to Equation~\ref{eq:desired_heading}.
\begin{equation} 
\label{eq:desired_heading} 
\text{Desired heading} = \frac{640\,\text{px}}{2d\tan{\frac{63^{\circ}}{2}}}{\times\frac{63^{\circ}}{640\,\text{px}}\times x}
\end{equation}
where $d$ is the distance between the ornithopter and the camera, and $x$ is 
the horizontal position of the robot relative to the center of the window in 
meters. This relation reflects the fact that the pose estimation algorithm 
does not determine the depth of the robot, which can be problematic for 
control algorithms. For example, if the ornithopter remains at a constant 
horizontal distance from the window, but moves closer and closer to the 
camera, the ground station's directed heading steadily increases, even 
though the robot has not actually moved farther from the window. These 
problems are most severe in areas close to the camera.

We model the pose estimation computation delay, which include the latency of 
updating the particle filter, calculating a new reference attitude, and 
sending the reference attitude command the ornithopter via radio.

@@

%............................................................................%
\subsubsection{Ornithoper Attitude Control}
We determined the turning speed and turning radius of the H$^2$Bird 
experimentally, by measuring the response of the system to a step input of a 
clockwise 90$^{\circ}$ turn. We measured the response of the system by 
recording the attitude estimates calculated by the ornithopter using its 
on-board gyroscope. We then used MATLAB\footnote{The MathWorks, Inc. Matlab:
\href{http://www.mathworks.com/products/matlab/}
     {http://www.mathworks.com/products/matlab/}} 
to fit a simple low-order model to the step response, as discussed in 
Sec.~\ref{sec:model_attitude}.

%............................................................................%
\subsubsection{Ground Station Pose Estimation}
To measure the capture latency of the camera, we wrote a simple test 
application which draws a large colored rectangle on the ground station's 
screen, then uses the camera to detect that changes in that rectangle's 
color. By noting the value of the system timer just before it changes the 
rectangle, and just after the camera detects that change, the application 
arrives at an estimate of the video capture pipeline latency. We averaged 
these result over several hundred rapidly-executed trials to determine the 
capture latency we used to model the system.

@@

%----------------------------------------------------------------------------%
\subsection{Model Verification}
\label{sec:experiments_verification}

We tested our cooperative visual servoing system with a simple target seeking
experiment: the H$^2$Bird must fly through a specified obstacle, in this 
case, a wooden frame we used to simulate a window.

\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{figures/experiment_cartoon.pdf}
\caption{Conceptual sketch of experimental setup.}
\label{fig:experiment_cartoon}
\end{figure}

As the H$^2$Bird cannot detect the target on its own, we used cooperative 
control with the ground station, which to provided remote guidance to 
H$^2$Bird to allow it to navigate to the window. We used a motion capture 
system to record ground truth data. We present a sketch of this experimental 
setup in Figure~\ref{fig:experiment_cartoon}.

Each trial included the following steps (see Fig.~\ref{fig:pf_screencap}):
\begin{enumerate}
\item We identify the window to the ground station by dragging a bounding 
box over it using the mouse. 
\item We release H$^2$Bird by hand, in view of the camera at the desired 
starting grid point.
\item The H$^2$Bird attempts to fly through the window with, with remote 
guidance from the ground station.
\end{enumerate}

After every 5 trials, we replaced the ornithopter's battery with a fully 
charged one.
